The Countercyclical Component of Capital Regulation

### 2020-04-14 Analysis/Draft/Draft_bookdown.Rmd
Remove RW decomposition of main analysis and improve results section

---
title: "Credit Risk Distribution and Capital Savings"
author:
- affiliation: CentER and EBC, Tilburg University; l.avezum@uvt.nl; +31134664531;  Warandelaan
    2, 5037 AB Tilburg, NL, The Netherlands
  name: Lucas Avezum
date: '\emph{\today}'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: pdflatex
    template: tex-ms.tex
  pdf_document: default
  word_document: default
biblio-style: apsr
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
csl: ../../Auxiliar/cambridge-journal-of-economics.csl
anonymous: no
fontsize: 11pt
geometry: margin= 1in
header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float}
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
linkcolor: black
abstract: The Basel II accord introduced a concave mapping from probabilities of default
  (PD) to capital requirements for banks adopting the internal ratings-based (IRB)
  approach. Therefore, total capital requirements depends not only on the portfolio's
  average PD but also on the shape of its PD distribution. Using a hand-collected
  dataset, I propose a method that decomposes the contribution of the PD distribution
  to capital requirements into an average component and a shape component. While a
  larger average component increases capital charges, a larger shape component generates
  capital savings. I document that these capital savings are important sources of
  risk-weight variability across banks. I further investigate the relationship of
  capital savings with the business cycle and find it to be countercyclical. The results
  suggest that a recalibration of the current IRB framework can significantly reduce
  the procyclicality of capital requirements.
spacing: double
subtitle: ''
thanks: Preliminary draft.
JEL: G21, G28, G11
urlcolor: black
toc: FALSE
---

@@

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

@@

# Institutional Background

@@

\begin{equation}
\label{measure1}
\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
\end{equation}

@@

\begin{equation}
\label{measure2}
\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
\end{equation}

@@

\begin{equation}
\label{measure3}
\text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
\end{equation}

@@

## RW Decomposition

@@

## A simple theoretical framework

@@

In Table \ref{tab1}, I regress the change in the logarithm of *Capital savings* on the change in the logatithm of real GDP per capita. In all regressions the measure of *Capital savings* follows equation \ref{eq4}. The difference between regressions in the table is the portfolio used to calculate the *Capital savings*. The real GDP per capita used  comes from the country where the bank has its headquarter. All regressions include bank and year fixed effects and standard errors clustered at the country level.

Column 1 shows that total *Capital savings* are countercyclical as the growth rate of tota l*Capital savings* is negatively associated to the GDP growth rate. According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average. The effect is economically sizeable representing 14% of total *Capital savings* standard deviation in the sample.

The remaining columns of Table \ref{tab1} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios. The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical. However, this coefficient is not statistically significant. On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant. According to the regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation. Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios. Only in regression 4 the coefficient of GDP growth rate is negative and statistically significant.

@@

A concern to the previous analysis is the possibility of that I could be overestimating the average impact of GDP growth rate on *Capital savings*. First, Because the measure used in Table \ref{tab1} is dimentionless, i.e., it is a measure of intensity of *Capital savings*, if the portfolio of small banks, in terms of asset, are more impacted during the business cycle than the portfolio of large banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank. To account for differences in *Capital savings* size, in Table \ref{tab2} the measure of *Capital savings* follows equation \ref{eq3}, which is a measure of *Capital savings* in monetary units.

The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.

Second, because both measures in Tables \ref{tab1} and \ref{tab2} are relative to the IRB portfolio, if the portfolio of banks with lower share of their portfolio under the IRB approach are more impacted during the business cycle than the portfolio of high IRB share banks then the cofficients in Table \ref{tab1} overestimate the impact to the average bank. The results shown in Table \ref{tab2} reinforce the evidence that the coutercyclicality of *Capital savings* is mostly driven by the corporate portfolio.

Lastly,  In Table \ref{tab3} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution. Again, the results shown in Table \ref{tab3} suppport the evidence from the last three tables: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.

## Capital Savings and Portfolio Realocation

Ideally we want to fix the porfolio at some point in time and compare its and new loans PD distribution. Unfortunately, I cannot track individual assets of a portfolio. In this section I compare moments of the PD distribution between good and bad times without disantangling the direct effect of the business cycle on the distribution from the realocation effect.

The results from Tables \ref{tab1} to XX raises the questions of why it appears that only corporate portfolios are impacted during the business cycle. 
I further analyse the relationship of the PD distribution moments with GDP and find that during economic expasions: average, variance and kurtosis decrease while the skewness increases. This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.

@@

```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r)%>%
  group_by(name, Country) %>%
#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
  group_by(Country) %>%
 # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
  #ungroup()%>%
  #mutate(ordering = order_aux+RWA.r,
  #       name = as.factor(name),
  #       name = fct_reorder(name, ordering),
  #       Country = fct_reorder(Country, ordering)) %>%
 ggplot(mapping = aes(x = reorder(name, RWA.r, FUN = mean, na.rm=TRUE),
                      y = RWA.r)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "RWA savings ratio") +
  scale_fill_grey()+
    theme_bw()

@@

```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
bank.data %>% filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  select(Country,name, year, RWA.r)%>%
  group_by(Country,name) %>%
 summarise_if(is.numeric,mean,na.rm= TRUE)%>%
 ggplot(mapping = aes(x = reorder(Country, RWA.r, FUN = mean, na.rm=TRUE),
                          y = RWA.r)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
  labs(y =  "RWA savings ratio") +
  coord_flip() +
  theme_bw()

@@

plot2 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year,  mean.PD) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=mean.PD)) +
  geom_point(alpha = 0.3) +
  geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = 'Average PD (%)', x=NULL)

plot3 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year, var.PD) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=sqrt(var.PD))) +
  geom_point(alpha = 0.3) +
  geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = 'SD PD (%)', x=NULL)

@@

```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
## Set winsorizing limits
p_w <- c(0.05, 0.95)        
bank.data.win <- bank.data %>% 
  ungroup()%>%
  filter(!is.na(RW.s), bvdid %in% pillar3.data$bvdid) %>% 
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  group_by(bvdid)

``` 
# Tables {-}

@@

# Variables in the model 
# Dependent variables
dep.vars <- c("", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
# Independent variables
ind.vars <- c("log.gdp_gr")
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")

@@

# Model formulas
formula.text <- paste0(paste0("log.RWA.r", dep.vars,"_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
   felm(x, data = filter(bank.data.win, RWA.s>0))
})

@@

```{r table_cyclicality_main, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          title = "Capital Savings and the Business Cycle",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} for portfolio $j$. In all regressions the measure of \\textit{Capital savings} follows equation \\ref{measure1}. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se               = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "small",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab1",
          style            = "qje",
          column.sep.width = "1pt",
          align            = TRUE)
```

\begin{table}[!htbp] \centering  
\caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
\label{tab2:reg}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{16cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline 
The dependent variable is the change in the logarithm of \textit{Capital savings} for portfolio $j$. In Panel A the measure of \textit{Capital savings} follows equation \ref{eq5}. In Panel B the measure of \textit{Capital savings} follows equation \ref{eq6}. In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}$} \\
\end{tabular}

@@

reg.list <- lapply(formulas, function(x) {
 felm(x, data = filter(bank.data.win, RWA.s>0))
})

@@

```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}

stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          float            = FALSE,
          align            = TRUE)

@@

stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          float            = FALSE,
          align            = TRUE)

@@

stargazer(reg.list,
          title = "Alternative measure: PD distribution Gini Coefficient",
          table.layout = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se               = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab3",
          style            = "qje",
          column.sep.width = "2pt",
          float            = FALSE, 
          align            = TRUE)

>>>>

---
title: "The Countercyclical Component of Capital Regulation"
date: '\emph{\today}'
output:
  bookdown::pdf_document2:
    fig_caption: yes
    latex_engine: pdflatex
    template: tex-ms.tex
  pdf_document: default
  word_document: default
author:
- name: "Lucas Avezum"
  affiliation: "CentER and EBC, Tilburg University"
biblio-style: apsr
bibliography: ../../Auxiliar/mybibfile.bib
citecolor: black
csl: ../../Auxiliar/cambridge-journal-of-economics.csl
#anonymous: no
fontsize: 11pt
geometry: margin= 1in
header-includes: \usepackage{rotating, graphicx, dcolumn, longtable, float, siunitx, tabularx}
keywords: Internal ratings-based models, Risk-weighted assets dispersion, Procyclicality
linkcolor: black
abstract: The Basel II accord introduced a concave mapping from probabilities of default
  (PD) to capital requirements for banks adopting the internal ratings-based (IRB)
  approach. Therefore, total capital requirements depends not only on the portfolio's
  average PD but also on the shape of its PD distribution. Using a hand-collected
  dataset, I propose a method that decomposes the contribution of the PD distribution
  to capital requirements into an average component and a shape component. While a
  larger average component increases capital charges, a larger shape component generates
  capital savings. I investigate the relationship of these capital savings with the business cycle and find it to   be countercyclical. I further document that the shape are important sources of across banks. The results suggest   that a recalibration of the current IRB framework can have sizeble effects on the procyclicality of capital       regulation.
spacing: double
subtitle: ''
thanks: Preliminary draft.
JEL: G21, G28, G11
urlcolor: black
toc: FALSE
---

@@

library(tidyverse)            ## Data manipulation, pipe operator
library(data.table)           ## Command to bind lists
library(ggridges)             ## Density ridges plot
library(stargazer)            ## Latex tables
library(lfe)                  ## High-dimensional fixed effects
library(gridExtra)            ## Create grid for plots and tables
library(cowplot)              ## Create grid for plots and tables
library(DescTools)            ## Command to winsorize variables
library(ggpubr)               ## More plots
library(splitstackshape)     ## Command to expand data

#library(textablr)
#library(sandwich)
#library(lmtest)
#library(relaimpo)            ## Calculate relative importance measures

@@

# Institutional Background and Capital Savings Measures

@@

\begin{equation}
\label{eq:measure1}
\text{RWA}^s=\frac{\text{RWA}^c}{\text{RWA}}.
\end{equation}

@@

\begin{equation}
\label{eq:measure2}
\text{RWA}^{\$}=\text{RWA}^c-\text{RWA}.
\end{equation}

Because this is a measure in US Dollars, it is affected by the size of the IRB portfolio. Therefore, differently from the main analysis, *Capital savings* are mechanically larger for larger banks and they increase as the portfolio rolls out to the IRB approach.

@@

\begin{equation}
\label{eq:measure3}
\text{RW}^s=\frac{\text{RWA}^{\$}}{\text{RWA}}.
\end{equation}

@@

@@

## A simple theoretical framework {#theory}

@@

Throughout this section, I report results from estimating the following specification:

\begin{equation*}
\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*}

The dependent variable is the yearly change in a variable associated to portfolio $j$ for bank $i$ at year $t$. The change in the logatithm of real GDP per capita used comes from the country where the bank has its headquarter. All regressions include bank and year fixed effects and robust standard errors adjusted for clustering at the country level are reported in parentheses.

Column 1 Table \ref{tab:main} shows results using equation \ref{eq:measure1}, my prefered measure of *Capital savings*.  shows that total *Capital savings* are countercyclical. According to regression 1 an increase of 1.6 percentage points in GDP growth rate (one standard deviation) decreases total *Capital savings* growth rate by 0.7 percentage points on average. The effect is economically sizeable representing 14% of total *Capital savings* standard deviation in the sample.

The remaining columns of Table \ref{tab:main} suggest that the countercyclical relationship between GDP and *Capital savings* is mostly driven by wholesale portfolios, especially corporate portfolios. The estimated coefficient in Column 2 is negative suggesting that *Capital savings* generated by retail portfolios are also countercyclical. However, this coefficient is not statistically significant. On the other hand, the coefficient of GDP growth rate on *Capital savings* from wholesale portfolios is statistically and economically significant. According to regression 3 one standard deviation increase in GDP growth rate decreases *Capital savings* from wholesale portfolios growth rate by 0.8 percentage points on average, which is 16% of the *Capital savings* from wholesale portfolios standard deviation. Columns 3 to 6 indicate that within the wholesale portfolio most of the countercyclical variation appears to come from corporate portfolios. Only in regression 4 the coefficient of GDP growth rate is negative and statistically significant.

@@

In this section, I test the robustness of the main results to alternative measures of *Capital savings*. In Panel A of Table \ref{tab:alt}, the dependent variable measures *Capital savings* in US dollars following equation \ref{eq:measure2}. As before, the estimated coefficients suggest that the countercyclical effect is mostly driven by wholesale portfolios. Differently from Table \ref{tab:main}, the coefficient of GDP growth rate estimated on the bank portfolio is now negative and statistically significant. Because coefficients in Panel A of Table \ref{tab:alt} are  impacted by the size of the portfolio, while in \ref{tab:main} they are not, the results of regressions 6 in both tables suggest that part of the *Capital savings* countercyclicality comes from an increase in exposure to other banks during recessions.

In panel B of Table \ref{tab:alt} the measure of *Capital savings* in US dollars is rescaled by the amount of EAD in the respective portfolio following equation \ref{eq:measure3}. This measure, therefore, defines *Capital savings* in terms of the usual metric of capital requirements intensity, risk-weights. Although the coefficient on the wholesale portfolio is no longer statiscally significant, total *Capital savings* and those coming from corporate portfolios are still statistically different from zero at the 1%  confidence level. The attenuation of the effect on wholesale portfolio appears to come from large, albeit not significant, procyclical *Capital savings* in sovereign portfolios. Moreover, the small and statistically insignificant coefficient in regression 6 reinforces the claim that countercyclical *Capital savings* in the bank portfolio are due to the extensive and not the intensive margin.

Lastly, in Panel C of Table \ref{tab:alt} I use the Gini coefficient as an alternative measure of the shape component of the PD distribution. Again, the results shown in Panel C suppport the evidence from Table \ref{tab:alt}: *Capital savings* are countercyclical and mostly driven by the effect of GDP growth rate on *Capital savings* from corporate portfolios.

## Capital Savings and Portfolio Reallocation

The results from Tables \ref{tab:main} and \ref{tab:alt} raises the question of why it appears that only *Capital savings* from corporate portfolios are impacted during the business cycle. In this section I argue that portfolio reallocation during the business cycle explains part of *Capital savings* cyclicality. Ideally, for this purpose, I would track individual assets during the cycle and compare their PDs evolution to the PDs of new loans.  Unfortunately, I can only observe assets aggregated by PD bands which does not allow me to track individual assets. As an alternative approach, I analyze the relationship of PD distribution moments with GDP to understand asset allocation during the business cycle.

According to the theoretical framework presented in section 4.1, an increase in average PD reduces the amount of *Capital savings* if other moments are kept constant. This relationship is illustrated in panel A of Figure \ref{fig:pdmapcases} and empirically tested in Table \ref{tab:moments_contribution}. In all columns of this table average PD is negatively associated with *Capital savings*. Table \ref{tab:moments_contribution} also shows that, in line with the theoretical framework, an increase in PD variance is positively associated with *Capital savings* for all portfolios. In sum, the estimated coefficients of the average PD and the PD variance provide strong empirical support to the theoretical framework presented in section 4.1. At the same time, however, I cannot refute the hypothesis that portfolio reallocation does not impact *Capital savings* because the estimated effect of these two moments on *Capital savings* are in line with a scenario without any portfolio reallocation.

Extending the analysis to the skewness and kurtosis of the PD distribution proves to be very helpful. Consider a

Panel A in table \ref{tab:moments_cyclicality} reports how average PD moves during the business cycle. As expected, in most of the regressions, changes in average PD are negatively associated with GDP growth. This association is the known driver of the procyclical character of the current model-based capital regulation and its interpretation is fairly mechanical. During recessions average credit quality deteriorates, or in other words, average PD increases. The only exception to the rule is the sovereign portfolio. Column 5 of panel A shows that the coefficient of GDP growth on changes on average PD of sovereign portfolios is positive, although without strong evidence that it is not different from zero. One possible reason for this result is that, during economic downturns, banks may search for safe assets mostly in the form of safe haven government bonds. According to column 5 of panel A, this reallocation of portfolio appears to be sufficient to counterbalance the deteriorations of the prior portfolio. It also helps explain the insignificant results in column 5 of Tables \ref{tab:main} and \ref{tab:alt}.

As explained, if no other moment is affected during the business cycle we should observe procyclical *Capital savings*. However, it was shown in

Common  
I further analyse the relationship of the PD distribution moments with GDP and find that during economic expansions: average, variance and kurtosis decrease while the skewness increases. This evidence is in line with the idea that the upper-tail (riskier borrowers) of the PD distribution is disproportionately affected by the business cycle.

@@

```{r, include = TRUE, fig.height = 6.5, fig.width = 5,results="hide",fig.cap= savings_cap, fig.align = 'center'}
bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid)  %>% 
  select(name, Country, RWA.r.Total)%>%
  group_by(name, Country) %>%
#   summarise_if(is.numeric,median,na.rm= TRUE)%>%
  group_by(Country) %>%
 # mutate(order_aux = 1000*mean(RWA.r,na.rm = TRUE))%>%
  #ungroup()%>%
  #mutate(ordering = order_aux+RWA.r,
  #       name = as.factor(name),
  #       name = fct_reorder(name, ordering),
  #       Country = fct_reorder(Country, ordering)) %>%
 ggplot(mapping = aes(x = reorder(name, RWA.r.Total, FUN = mean, na.rm=TRUE),
                      y = RWA.r.Total-1)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
    coord_flip() +
  labs(y =  "RWA savings ratio") +
  scale_fill_grey()+
    theme_bw()

@@

```{r, include=TRUE, fig.height = 4, fig.width = 3,results="hide",fig.cap= savings_cap_country, fig.align = 'center'}
bank.data %>% filter(RWA.r.Total>0, bvdid %in% pillar3.data$bvdid) %>% 
  select(Country,name, year, RWA.r.Total)%>%
  group_by(Country,name) %>%
 summarise_if(is.numeric,mean,na.rm= TRUE)%>%
 ggplot(mapping = aes(x = reorder(Country, RWA.r.Total, FUN = mean, na.rm=TRUE),
                          y = RWA.r.Total-1)) +
geom_point(alpha = 0.5,size = 2) +
  stat_summary(fun.y="mean", geom="bar",alpha = 0.3) +
    labs(x=NULL) +
  labs(y =  "RWA savings ratio") +
  coord_flip() +
  theme_bw()

@@

plot2 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year,  mean.PD.Total) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=mean.PD.Total)) +
  geom_point(alpha = 0.3) +
  geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = 'Average PD (%)', x=NULL)

plot3 <-bank.data %>% filter(year>2007)%>%
  select(bvdid, year, var.PD.Total) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))%>%
  ggplot(mapping = aes(x = year,y=sqrt(var.PD.Total))) +
  geom_point(alpha = 0.3) +
  geom_smooth(stat = 'summary',
            fun.data = mean_cl_quantile,
            color = "black")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = 'SD PD (%)', x=NULL)

@@

```{r winsorized_data, fig.height = 10, fig.width = 7, results="asis",out.width = "1\\textwidth"}
## Set winsorizing limits
p_w <- c(0.05, 0.95)        
bank.data.win <- bank.data %>% 
  ungroup()%>%
  mutate_if(is.numeric,Winsorize,probs = p_w ,na.rm = TRUE)%>%
  group_by(bvdid)

```{r PD_distribution_data, eval=FALSE, echo=FALSE}

IRB.data <- as.data.frame(pillar3.data) %>%
  filter(Method2 %in% c("IRB"),
         PD  < 0.99, EAD > 0) %>%
  mutate(LGD        = ifelse(is.na(LGD), 0.45, LGD),
         w.PD       = PD*EAD*LGD,
         w.EAD      = LGD*EAD)

PD.distribution <- data.frame()

for(i in c("Total","Wholesale","Retail","Corporate","Sovereign","Banks")){
PD.distribution <- IRB.data %>%
  select(bvdid, name, year, w.EAD, PD, Portfolio_1, Portfolio_2) %>%
  {if(i %in% c("Total")){
     .
    }else if(i %in% c("Wholesale","Retail")){
     filter(., Portfolio_1 %in% c(i))
    }else{
     filter(., Portfolio_2 %in% c(i))}} %>%
  mutate(w.EAD = round(w.EAD/100)) %>%
  na.omit(cols=c("w.EAD")) %>%
  expandRows("w.EAD") %>%
  rbind(PD.distribution)}

```{r}
density_cap = "\\label{fig:pdmapcases}The figure shows the two possible outcomes of a negative credit risk shock on a portfolio. The left-hand side plot shows the case where capital savings are procyclical. The right-hand side plot shows the case where capital savings are countercyclical. "
```

```{r, density, eval=FALSE, fig.height = 3, fig.width = 6,results="hide",out.width = "90%", fig.align = 'center', echo=FALSE, message=FALSE, fig.cap= denisty_cap}

ggplot(PD.distribution) + geom_density(aes(x=PD))

view(select(pillar3.data, name, year, PD, Portfolio_1, Portfolio_2))
```

# Tables {-}

@@

# Variables in the model 
# Dependent variables
dep.vars <- c(".Total", ".Retail", ".Wholesale", ".Corporate", ".Sovereign", ".Banks")
dep.text <- c("$j = \\text{Total}$","$j = \\text{Retail}$", "$j = \\text{Wholesale}$", "$j = \\text{Corporate}$", "$j = \\text{Sovereign}$", "$j = \\text{Banks}$")
# Independent variables
ind.vars <- c("log.gdp_gr")
ind.text <- c("$\\Delta\\text{Log GDP}_{i,t}$")

@@

# Model formulas
formula.text <- paste0(paste0("log.RWA.r", dep.vars, "_gr")," ~ ", paste0(ind.vars)," | year + bvdid | 0 | Country ")
formulas <- lapply(formula.text, formula)

# Run the models
reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win)
})

@@

```{r table_cyclicality_main, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
stargazer(reg.list,
          title = "Capital Savings and the Business Cycle",
          table.layout = "n=c=d#-t-as-",
          notes = "\\parbox[t]{16.5cm}{The table shows estimates for the following model: \\newline \\begin{equation*}\\Delta \\text{Log RWA}_{i,t}^{s,j} = \\beta \\Delta\\text{Log GDP}_{i,t}+\\alpha_{i}+\\alpha_{t}+\\varepsilon_{i,t}. \\end{equation*} \\newline The dependent variable is the change in the logarithm of \\textit{Capital savings} from portfolio $j$. In all regressions the measure of \\textit{Capital savings} follows equation \\ref{eq:measure1}. $\\Delta\\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\\%, 5\\%, and 1\\% level respectively.}",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se               = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          label            = "tab:main",
          style            = "qje",
          column.sep.width = "1pt",
          align            = TRUE)
```

\begin{table}[!htbp] \centering  
\caption{Capital Savings and the Business Cycle: Robustness to Alternative Measures}  
\label{tab:alt}
\footnotesize
\begin{tabular}{@{\extracolsep{1pt}}lccccc}
\multicolumn{6}{c}{\parbox[t]{15.8cm}{The table shows estimates for the following model: \newline \begin{equation*}\Delta \text{Y}_{i,t}^{j} = \beta \Delta\text{Log GDP}_{i,t}+\alpha_{i}+\alpha_{t}+\varepsilon_{i,t}. \end{equation*} \newline 
The dependent variable is the change in the logarithm of \textit{Capital savings} from portfolio $j$. The measure of \textit{Capital savings} follows equation \ref{eq:measure2} in panel A and equation \ref{eq:measure3} in panel B. In Panel C the Gini coefficient is used as alternative measure of the shape of the PD distribution. $\Delta\text{Log GDP}_{i,t}$ is the real GDP per capita growth rate. All regressions include bank and year fixed effects. Robust standard errors adjusted for clustering at the country level are reported in parentheses. *, **, and *** indicate statistical significance at the 10\%, 5\%, and 1\% level respectively.}} \\
[-1.8ex] \\
\hline \\
\end{tabular}\\
\begin{tabular}{@{\extracolsep{1pt}}cccccc}
\multicolumn{6}{c}{Panel A: $\text{Y}=\text{Log RWA}^{\$}$} \\
\end{tabular}

@@

reg.list <- lapply(formulas, function(x) {
   felm(x, data = bank.data.win)
})

@@

```{r table_cyclicality_alt, eval=TRUE, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          #label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

@@

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          #label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

@@

table <- capture.output(
{stargazer(reg.list,
          table.layout     = "=c=d#-t-as-",
          dep.var.labels   = dep.text,
          covariate.labels = ind.text,
          #se              = se.list,
          add.lines        = fe.list,
          keep.stat        = c("n","rsq"),
          header           = FALSE, 
          type             = "latex",
          font.size        = "footnotesize",
          notes.label      = "",
          notes.align      = "c", 
          notes.append     = FALSE,
          #label            = "tab2",
          style            = "qje",
          column.sep.width = "2pt",
          align            = TRUE,
          float            = FALSE)})

table <- gsub("{-3}",
              "{5.5}"
              , table, fixed = T)

cat(table)

```
\end{table}

```{r reg_cyclicality_components, echo=FALSE}
